---
title: "Multiple regression - beaver biodiv"
author: Liz Anderson
format: html
---

Copied over work from previous qmd (`high_prob_beaver_biodiv.qmd`)



## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```

```{r}
## set up the data frames

results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog")

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)

## land use data
lulc <- read.csv(here("Data", "LULC_spillover_diversity.csv"))
glimpse(lulc)
```


To prep data for multiple regression need to:

Columns for SR, treatment, site, perc developed, perc other habitat types

```{r}
results_wild_wide <- results_wild |>
  group_by(treatment, deployment_num, species) |>
  summarise(num_detections = n()) |>
  pivot_wider(names_from = species, values_from = num_detections)


results_wild_wide[is.na(results_wild_wide)] <- 0

results_wild_vegan <- results_wild_wide |>
  ungroup() |>
  select(-treatment, -deployment_num)

Div_Indices <- results_wild_wide[,1:2]
Div_Indices$SR <- specnumber(results_wild_vegan) 
Div_Indices

## don't need treatment here because will be joining with dataset that already has treatment

Div_Indices_SR <- Div_Indices |>
  ungroup() |>
  select(-treatment)
```


## Landcover data


To do:
* Convert to long
* filter to just the percentages

```{r}
## just the percentage variables
lulc_narrow <- lulc |>
  select(1, 2, 17:29) 
glimpse(lulc_narrow)
glimpse(lulc)

lulc_combined <- lulc_narrow |>
  mutate(p_Developed = p_Developed_open + p_Developed_low_intensity + p_Developed_med_intensity) |>
  mutate(p_water = p_Open_water + p_Woody_wetlands + p_Emergent_herb_wetlands) |>
  mutate(p_forest = p_Deciduous + p_Evergreen + p_Mixed_forest) |>
  mutate(p_grassy = p_Grassland_herbaceous + p_Pasture_hay + p_Shrub_scrub) |>
  relocate(p_Developed, p_water, p_forest, p_grassy, .after = area_ha) |>
  mutate(treatment = if_else(DEPLOYMENT_NUM == "AM0067" | DEPLOYMENT_NUM == "AM0066" | DEPLOYMENT_NUM == "AM0086" | DEPLOYMENT_NUM == "AM0089" | DEPLOYMENT_NUM == "AM0075" | DEPLOYMENT_NUM == "AM0073" | DEPLOYMENT_NUM == "AM0074" | DEPLOYMENT_NUM == "AM0090" | DEPLOYMENT_NUM == "AM0065",
                             true = "control",
                             false = "beaver")) |>
  relocate(treatment)
  
```


```{r}
## join SR with lulc data

lulc_sr_all <- left_join(Div_Indices_SR, lulc_combined, by = join_by(deployment_num == DEPLOYMENT_NUM))

lulc_sr_7.7 <- lulc_sr_all |>
  filter(area_ha == "7.7")

lulc_sr_17.6 <- lulc_sr_all |>
  filter(area_ha == "17.6")

lulc_sr_1.1 <- lulc_sr_all |>
  filter(area_ha == "1.1")
```


```{r}
## explore variables to see what predictors to include

ggplot(data = lulc_sr_7.7, aes(x = treatment, y = p_water)) +
  geom_boxplot()
## a bit more in beaver but not much

ggplot(data = lulc_sr_7.7, aes(x = treatment, y = p_Developed)) +
  geom_boxplot()
## more in control

ggplot(data = lulc_sr_7.7, aes(x = treatment, y = p_forest)) +
  geom_boxplot()
## about the same means

ggplot(data = lulc_sr_7.7, aes(x = treatment, y = p_grassy)) +
  geom_boxplot()
## more in beaver

ggplot(data = lulc_sr_7.7, aes(x = treatment, y = p_Woody_wetlands)) +
  geom_boxplot()
## p_woody_wetlands about same
## p_emergent_wetlands more in beaver none in control

ggplot(data = lulc_sr_7.7, aes(x = treatment, y = p_Open_water)) +
  geom_boxplot()
## not much open water in either with outlier exception in beaver
```


Add in the active vs not active site data
```{r}
beaver_active_df <- tibble(deployment_num = c("AM0063", "AM0064", "AM0081", "AM0083", "AM0087", "AM0094", "AM0095", "AM0096", "AM0067", "AM0066", "AM0086", "AM0089", "AM0075", "AM0073", "AM0074", "AM0090"), 
       active_yn = c("hist", "hist", "yes", "yes", "yes", "yes", "yes", "hist", "no", "no", "no", "no", "no", "no", "no", "no"))

## paused here because not sure how to distinguish past presence with 
lulc_sr_7.7 <- left_join(lulc_sr_7.7, beaver_active_df, by = join_by(deployment_num))

lulc_sr_7.7 <- lulc_sr_7.7 |>
  relocate(active_yn)

lulc_sr_17.6 <-left_join(lulc_sr_17.6, beaver_active_df, by = join_by(deployment_num)) |>
  relocate(active_yn)
```


### multiple regression
Mult regression model 1 --
beaver predictors 
* treatment
* active
* p_water
* p_grassy

human predictors
* p_developed

```{r}
beaver_model <- lm(data = lulc_sr_7.7, SR ~ treatment + p_water + p_grassy) # activy_yn
summary(beaver_model)

human_model <- lm(data = lulc_sr_7.7, SR ~ p_Developed + treatment)
summary(human_model)


## full model
full_model_short <- lm(data = lulc_sr_7.7, SR ~ p_Developed + treatment + p_water)
summary(full_model_short)
```

## 17.6 area models 
### full model then stepwise regression

```{r}
full_model <- lm(data = lulc_sr_17.6, SR ~ treatment + p_Developed + p_water + p_forest + p_grassy)
summary(full_model)


backward_model <- step(full_model, direction = "backward")
## backward doesn't take anything out

intercept_only <- lm(SR ~ 1, data = lulc_sr_17.6)

forward_model <- step(intercept_only, direction = "forward", scope = formula(full_model))
## best: treatment only

both_model <- step(intercept_only, direction = "both", scope = list(lower = intercept_only, upper = full_model))
## treatment only (had open_water in there before took out)
```

```{r}
old_best_step_both <- lm(SR ~ treatment + p_Open_water, data = lulc_sr_17.6)
summary(old_best_step_both)

best_step_both_17.6 <- lm(SR ~ treatment, data = lulc_sr_17.6)
summary(best_step_both_17.6)
```

## 7.7 area level
```{r}
full_model_7.7 <- lm(data = lulc_sr_7.7, SR ~ treatment + p_Developed + p_water + p_forest + p_grassy)
summary(full_model_7.7)


backward_model_7.7 <- step(full_model_7.7, direction = "backward")
## best: SR ~ p_Developed + p_water + p_forest

intercept_only_7.7 <- lm(SR ~ 1, data = lulc_sr_7.7)

forward_model_7.7 <- step(intercept_only_7.7, direction = "forward", scope = formula(full_model_7.7))
## best: SR ~ p_Developed + p_grassy

both_model_7.7 <- step(intercept_only_7.7, direction = "both", scope = list(lower = intercept_only_7.7, upper = full_model_7.7))
## best: SR ~ p_Developed + p_grassy
```

```{r}
old_best_7.7 <- lm(SR ~ p_Developed + p_Open_water + treatment, data = lulc_sr_17.6)
summary(old_best_7.7) ## not great

best_7.7 <- lm(SR ~ p_Developed + p_grassy, data = lulc_sr_17.6)
summary(best_7.7)
```


## GLM models

```{r}
## want to have control be the baseline - need to relevel
lulc_sr_7.7 <- lulc_sr_7.7 |>
  mutate(treatment = as.factor(treatment)) |>
  mutate(treatment = relevel(treatment, "control"))
```


```{r}
glm_mod_full <- glm(SR ~ treatment + p_Developed + p_water + p_forest + p_grassy, data = lulc_sr_7.7, family = "poisson")
summary(glm_mod_full)
## having p_grassy gives warning that did not converge (I think b/c no diff)
```

```{r}
glm_mod_limited <- glm(SR ~ treatment + p_Developed + p_water + p_forest, data = lulc_sr_7.7, family = "poisson")
summary(glm_mod_limited)
```

```{r}
glm_mod_lulc_only <- glm(SR ~ p_Developed + p_water + p_forest, data = lulc_sr_7.7, family = "poisson")
summary(glm_mod_lulc_only)
```

```{r}
glm_mod_threePred <- glm(SR ~ treatment + p_Developed + p_water, data = lulc_sr_7.7, family = "poisson")
summary(glm_mod_threePred)
```

#### Try stepwise with glm
```{r}
step(glm_mod_limited)
```

#### Test for overdispersion 
(according to google) divide residual deviance by degrees of freedom. If much over 1, then overdispersed. 
Can also use the following package and function to check:
```{r}
library(performance)
check_overdispersion(glm_mod_limited)
## no overdispersion detected. (So poisson model works)
```

```{r}
best_mod_7.7 <- glm(formula = SR ~ p_Developed + p_water + p_forest, family = "poisson", 
    data = lulc_sr_7.7)

summary(best_mod_7.7)
```

#### Check for multicolinearity
```{r}
library(car)
vif(best_mod_7.7)
## above 5 or 10 means problematic multicolinearity
```

#### Reduce multicolinearity
```{r}
mod_less_colin <- glm(SR ~ p_Developed + p_water, family = "poisson", 
    data = lulc_sr_7.7)
summary(mod_less_colin)

vif(mod_less_colin) ## not problematic multicolinearity
```

```{r}
## test with 17.6 data and same model
mod_less_colin_17.6 <- glm(SR ~ p_Developed + p_water, family = "poisson", data = lulc_sr_17.6)
summary(mod_less_colin_17.6) ## not signif at this level

vif(mod_less_colin_17.6) # no multicolin problem


## check with 1.1 data and same model
mod_less_colin_1.1 <- glm(SR ~ p_Developed + p_water, family = "poisson", data = lulc_sr_1.1)
summary(mod_less_colin_1.1) ## also not signif at this level

vif(mod_less_colin_1.1) ## no mulicolin

```





## Descriptive stats of lulc data

```{r}
habitat_means <- lulc_sr_7.7 |>
  group_by(treatment) |>
  summarise(p_development = mean(p_Developed),
            p_water = mean(p_water),
            p_forest = mean(p_forest),
            p_grassy = mean(p_grassy))

habitat_sd <- lulc_sr_7.7 |>
  group_by(treatment) |>
  summarise(p_development = sd(p_Developed),
            p_water = sd(p_water),
            p_forest = sd(p_forest),
            p_grassy = sd(p_grassy))

habi_means <- habitat_means |>
  pivot_longer(cols = 2:5, names_to = "habitat", values_to = "means")

habi_sd <- habitat_sd |>
  pivot_longer(cols = 2:5, names_to = "habitat", values_to = "sd")

habi_joined <- left_join(habi_means, habi_sd, by = join_by(habitat, treatment))
```

```{r}
plot_habi_means <- ggplot(data = habi_joined, aes(x = treatment, y = means)) +
  geom_errorbar(data = habi_joined, aes(ymin = means - sd, ymax = means + sd, color = treatment), width = 0.2, linewidth = 1) +
  geom_point() +
  facet_wrap(~habitat) +
  theme_bw() +
  scale_color_manual(values = c("#2c7fb8", "#7fcdbb")) +
  labs(x = "Treatment",
       y = "Mean percent habitat",
       color = "Treatment")

ggsave(here("Output", "plot_habi_means.png"), plot = plot_habi_means, width = 7, height = 5)
```



